On the number of tiles visited by a\\ line segment on a rectangular grid

### 2021-09-18 paper.tex
Added references on Buffon's needle problem
Typos

\newcommand{\mr}{\mathrm}
\newcommand{\st}{\mid}%{\,|\,}
\newcommand{\cond}{\,|\,} % less spacing that \mid
\newcommand{\diff}{\,\mathrm d}
\let\Re\relax % remove the command...
\DeclareMathOperator{\Re}{Re} % ...to define it differently
\let\Im\relax
\DeclareMathOperator{\Im}{Im}
\DeclareMathOperator{\E}{E}
\newcommand{\funt}{\tau} % function that gives maximum number of touched tiles for real-valued length
\newcommand{\funl}{\lambda} % function that gives infimum real-valued length for a given number of touched tiles
\newcommand{\funti}{T} % function that gives maximum number of touched tiles for integer-valued length
\newcommand{\funli}{\Lambda} % function that gives minimum integer-valued length for a given number of touched tiles
\newcommand{\funta}{\varphi} % function that gives average number of touched tiles for real-valued length
\newcommand{\probmax}{\rho} % function that gives the probability of achieving the maximum number of touched files in the square case for real-valued length
\newcommand{\len}{\ell} % length variable, real-valued
\newcommand{\leni}{\ell} % length variable, integer-valued. Same notation, but I keep the different command just in case
\newcommand{\tiles}{t} % number of tiles (integer-valued)
\newcommand{\isolr}{i^+}
\newcommand{\jsolr}{j^+}
\newcommand{\isoli}{i^\ast}
\newcommand{\jsoli}{j^\ast}
\newcommand{\genfun}{q}
\newcommand{\genvar}{r}
\newcommand{\mss}{M}

@@

The problems studied in this paper are related to \ldots ***maybe a short paragraph here mentioning related problems: Buffon's needle, Buffon-Laplace problem. Some references: ``Buffon's Noodle Problem'', J.F.~Ramaley, 1969. ``On Laplace's Extension of the Buffon Needle Problem'', B.J.~Arnow, 1994. \emph{An Introduction to Geometrical Probability}, A.M~Mathai, section 1.1. \\
\url{https://www.jstor.org/stable/2317945} \\
\url{https://www.jstor.org/stable/2687085}

@@

\begin{proof}
% *** Is this writeup too informal/non-rigorous?
% ***LM: I would make the rational-length extension more explicit, and invoke a suitable theorem (with the corresponding book reference) for the real-length case. Or better yet: the first reference linked as a comment in the introduction (https://www.jstor.org/stable/2317945?seq=1#metadata_info_tab_contents) already gives the $2\len/(\pi a)$ formula. We can directly start from that formula, referencing that paper.
The classic Buffon's Needle result (**AA cite Buffon's 1777 paper?** **LM I think maybe something more modern such as the references that are linked in the Introduction section**) states that if a needle of length $1$ is tossed with uniformly random position and orientation on
infinite ruled paper with parallel lines a distance of $1$ apart the probability that it crosses any line is $2/\pi$.

@@

There are two situations depending on whether $\tiles$ is odd or even. Consider $\tiles$ odd first. Under this assumption, the range of lengths corresponding to $\tiles$ is given by \eqref{eq: square, odd max tiles, lengths}. For any length in this interval, there are some positions within the reference file from which the tile with lower-left coordinate $((\tiles-1)/2, (\tiles-1)/2)$ can be reached. Equivalently, the shifted segment with one endpoint at $((\tiles-1)/2, (\tiles-1)/2)$ can reach the reference tile. This is illustrated in Figure~\ref{fig: probmax_funt_odd_diag}. In addition, if 
\begin{equation}
\label{eq: probmax len odd2}
\len > \frac{\sqrt{(\tiles-5)^2 + (\tiles-1)^2}}{2}
\end{equation}
the two tiles with lower-left coordinates $((\tiles-3)/2, (\tiles+1)/2)$ and $((\tiles+1)/2, (\tiles-3)/2)$ are also reachable from the reference tile; or equivalently the shifted segment can reach the reference tile, as shown in Figure~\ref{fig: probmax_funt_odd_offdiag}. This can only happen for $\tiles \geq 5$ (for $\tiles=3$ the inequality \eqref{eq: probmax len odd2} cannot be satisfied for lengths in the range given by \eqref{eq: square, odd max tiles, lengths}). In this case, by symmetry, each of those two tiles contributes the same probability. Other tiles corresponding to the same $\tiles$ need not be considered, because they cannot be reached with lengths satisfying \eqref{eq: square, odd max tiles, lengths}.

>>>>

\newcommand{\mr}{\mathrm}
\newcommand{\st}{\mid}%{\,|\,}
\newcommand{\cond}{\,|\,} % less spacing that \mid
\newcommand{\diff}{\,\mathrm d}
\let\Re\relax % remove the command...
\DeclareMathOperator{\Re}{Re} % ...to define it differently
\let\Im\relax
\DeclareMathOperator{\Im}{Im}
\DeclareMathOperator{\E}{E}
\newcommand{\funt}{\tau} % function that gives maximum number of touched tiles for real-valued length
\newcommand{\funl}{\lambda} % function that gives infimum real-valued length for a given number of touched tiles
\newcommand{\funti}{T} % function that gives maximum number of touched tiles for integer-valued length
\newcommand{\funli}{\Lambda} % function that gives minimum integer-valued length for a given number of touched tiles
\newcommand{\funta}{\varphi} % function that gives average number of touched tiles for real-valued length
\newcommand{\probmax}{\rho} % function that gives the probability of achieving the maximum number of touched tiles in the square case for real-valued length
\newcommand{\len}{\ell} % length variable, real-valued
\newcommand{\leni}{\ell} % length variable, integer-valued. Same notation, but I keep the different command just in case
\newcommand{\tiles}{t} % number of tiles (integer-valued)
\newcommand{\isolr}{i^+}
\newcommand{\jsolr}{j^+}
\newcommand{\isoli}{i^\ast}
\newcommand{\jsoli}{j^\ast}
\newcommand{\genfun}{q}
\newcommand{\genvar}{r}
\newcommand{\mss}{M}

@@

The problems studied in this paper are related to \ldots ***maybe a short paragraph here mentioning related problems: Buffon's needle, Buffon-Laplace problem. Some references: \cite{Ramaley69}, \cite{Arnow94}, \cite{Mathai99}.
%\url{https://www.jstor.org/stable/2317945}
%\url{https://www.jstor.org/stable/2687085}

@@

\begin{proof}
% ***AA: Is this writeup too informal/non-rigorous?
% ***LM: \cite{Ramaley69} already gives the $2\len/(\pi a)$ formula. We can directly start from that formula, referencing that paper, which simplies the proof.
The classic Buffon's Needle result (**AA cite Buffon's 1777 paper?** **LM I think maybe something more modern such as the references that are linked in the Introduction section**) states that if a needle of length $1$ is tossed with uniformly random position and orientation on
infinite ruled paper with parallel lines a distance of $1$ apart the probability that it crosses any line is $2/\pi$.

@@

There are two situations depending on whether $\tiles$ is odd or even. Consider $\tiles$ odd first. Under this assumption, the range of lengths corresponding to $\tiles$ is given by \eqref{eq: square, odd max tiles, lengths}. For any length in this interval, there are some positions within the reference tile from which the tile with lower-left coordinate $((\tiles-1)/2, (\tiles-1)/2)$ can be reached. Equivalently, the shifted segment with one endpoint at $((\tiles-1)/2, (\tiles-1)/2)$ can reach the reference tile. This is illustrated in Figure~\ref{fig: probmax_funt_odd_diag}. In addition, if 
\begin{equation}
\label{eq: probmax len odd2}
\len > \frac{\sqrt{(\tiles-5)^2 + (\tiles-1)^2}}{2}
\end{equation}
the two tiles with lower-left coordinates $((\tiles-3)/2, (\tiles+1)/2)$ and $((\tiles+1)/2, (\tiles-3)/2)$ are also reachable from the reference tile; or equivalently the shifted segment can reach the reference tile, as shown in Figure~\ref{fig: probmax_funt_odd_offdiag}. This can only happen for $\tiles \geq 5$ (for $\tiles=3$ the inequality \eqref{eq: probmax len odd2} cannot be satisfied for lengths in the range given by \eqref{eq: square, odd max tiles, lengths}). In this case, by symmetry, each of those two tiles contributes the same probability. Other tiles corresponding to the same $\tiles$ need not be considered, because they cannot be reached with lengths satisfying \eqref{eq: square, odd max tiles, lengths}.
